---
title: "Final Project"
author: "Mariana MacDonald"
date: '2022-05-22'
output:
  pdf_document: default
bibliography: bibliography.bib
---

## Data importing and cleaning steps are explained in the text and follow a logical process. Outline your data preparation and cleansing steps.

- familiarized with the data sets;
- checked for NAs, errors or missing values;
- changed the names of the variables when needed to make it standard and easier to read and use;
- extracted only relevant variables from the data sets for my research;
- most of the data I am using, is already clean



## With a clean dataset, show what the final data set looks like. However, do not print off a data frame with 200+ rows; show me the data in the most condensed form possible.

DATA1 - Mental health Depression disorder Data 

```{r}
setwd("/Users/marianamacdonald/Documents/DATA SCIENCE/DSC 520/Statistics R/Week 2/dsc520")
library(readxl)
library(dplyr)
DATA1 <- read_excel("DATA 1 - Final Project - Mental health Depression disorder Data.xlsx")
library(tidyverse)
library(magrittr)
library(tidyr)
library(purrr)
names(DATA1)

colnames(DATA1) <- c("entity", "code", "year", "schizophrenia", "bipolar_disorder",
                     "eating_disorders","anxiety_disorders", "drug_use_disorders", "depression",
                     "alcohol_use_disorders")

newdata1 <- subset(DATA1, code == "USA", select=c(code, year, depression))
head(newdata1)
```

```{r}
newdata2<- subset(DATA1, select=c(code, depression))
head(newdata2)
```





DATA 2 - Glassdoor Gender Pay Gap


```{r}
setwd("/Users/marianamacdonald/Documents/DATA SCIENCE/DSC 520/Statistics R/Week 2/dsc520")
DATA2 <- read.csv("DATA 2 - Final Project - Glassdoor Gender Pay Gap.csv", header = T,
                  stringsAsFactors = T)
names(DATA2)
head(DATA2)
```



DATA 3 - The Depression Dataset -

From this data, I intend to use only a few variables, I have removed the melanch and inpatient, which had NA values and I am not interested on them. Also, from the entire dataset, I am only able to use  a few rows. The others are missing basically all the information. Conditions 7,8,9 have been removed.

```{r}
library(tidyr)
 
setwd("/Users/marianamacdonald/Documents/DATA SCIENCE/DSC 520/Statistics R/Week 2/dsc520")
DATA3 <- read.csv("DATA 3 - Final Project - scores.csv", header = T)
names(DATA3)

newdata3 <- DATA3 %>% drop_na(afftype, melanch)
newdata3
```

Description of variables
number (patient identifier), days (number of days of measurements), gender (1 or 2 for female or male), age (age in age groups), afftype (1: bipolar II, 2: unipolar depressive, 3: bipolar I), melanch (1: melancholia, 2: no melancholia), inpatient (1: inpatient, 2: outpatient), edu (education grouped in years), marriage (1: married or cohabiting, 2: single), work (1: working or studying, 2: unemployed/sick leave/pension), madrs1 (MADRS score when measurement started), madrs2 (MADRS when measurement stopped).





## What do you not know how to do right now that you need to learn to import and cleanup your dataset?

I have learned how to import csv, excel and arff dataset so I believe I have learned what I need for this project. What is pending is machine learning.


## Discuss how you plan to uncover new information in the data that is not self-evident.

At this moment, I am not sure if the predictors I am selecting will have relationship to the questions I want to answer, so I might need to use other variables to get to my solution. I might use correlation, regression, ANOVA, histograms and/or graphs to uncover new information.

## What are different ways you could look at this data to answer the questions you want to answer?

DATA1
Instead of only considering depression, I can sum the % of all the mental disorders and create a new variable (called Sum)

```{r}
library(readxl)
getwd()
disorders_df <- read_excel("DATA 1 - Final Project - Mental health Depression disorder Data.xlsx")
head(disorders_df)
colnames(disorders_df)
```
```{r}
disorders_df$Sum <- rowSums(disorders_df[c('Schizophrenia (%)', 'Bipolar disorder (%)',
'Eating disorders (%)', 'Anxiety disorders (%)',
'Drug use disorders (%)', 'Depression (%)',
'Alcohol use disorders (%)')], na.rm = TRUE)
head(disorders_df)
```


DATA 2
I can separate the data into male and female base pay, and look at the summary to find the mean and compare. (Male USD 98,458 x Female USD 89,943)

```{r}
malepay <- subset(DATA2, Gender == "Male", select=c(Gender, BasePay))
head(malepay)
summary(malepay)

femalepay <- subset(DATA2, Gender == "Female", select=c(Gender, BasePay))
head(femalepay)
summary(femalepay)
```



## Do you plan to slice and dice the data in different ways, create new variables, or join separate data frames to create new summary information? Explain.

I will not be joining data frames. They are very different and I won't benefit from joining them. I might create new variables.

## How could you summarize your data to answer key questions?

```{r}
summary(newdata1)
```


```{r}
summary(newdata2)
```


```{r}
summary(DATA2)
```

```{r}
summary(newdata3)
```



## What types of plots and tables will help you to illustrate the findings to your questions? Ensure that all graph plots have axis titles, legend if necessary, scales are appropriate, appropriate geoms used, etc.).


DATA 1

Depression in the USA during the years

```{r}
library(ggplot2)
theme_set(theme_minimal())
ggplot(newdata1, aes(x=year, y=depression)) + geom_point() + 
  ggtitle("Depression in the USA")+xlab("Country)") + ylab("Depression %)")
```

Table of USA, Year and Depression %

```{r table, echo=FALSE, results='asis'}
library(pander)
pandoc.table(newdata1)
```


DATA 2

gap pay between male and female

```{r}
qplot(Gender, BasePay, geom = "boxplot", data = DATA2, na.rm=TRUE, fill=Gender)
```

Education x gender

```{r}
library(dplyr)
DATA2 %>% group_by(Gender, Education) %>% summarize(count = n())
```

```{r}
library(dplyr)
new_glassdoor_df <- DATA2 %>% group_by(Gender, Education) %>% summarize(count = n())

library(ggplot2)
ggplot(new_glassdoor_df, aes(Education, count, fill = Gender)) +
geom_bar(stat="identity", position = 'dodge') +
labs(title="Multiple Bar plots")
```

DATA 3 

Depression by gender 

Regrettably, this is not a very good data to analyze this correlation. Based on many studies that I will discuss at the next step, about twice as many women as men experience depression [@Mayo]


"The prevalence of major depression is higher in women than in men;6,7 in 2010 its global annual prevalence was 5.5% and 3.2%, respectively, representing a 1.7-fold greater incidence in women." [@ncbi]


```{r}

library(ggplot2)

newdata3[newdata3$gender==1, 'gender'] <- "female"
newdata3[newdata3$gender==2, 'gender'] <- "male"

newdata3[newdata3$afftype==1, 'afftype'] <- "bipolar II"
newdata3[newdata3$afftype==2, 'afftype'] <- "depression"
newdata3[newdata3$afftype==3, 'afftype'] <- "bipolar I"

ggplot(newdata3, 
       aes(x = gender, 
           fill = afftype)) + 
  geom_bar(position = position_dodge(preserve = "single"))
```

Marriage x Depression

```{r}

library(ggplot2)

newdata3[newdata3$marriage==1, 'marriage'] <- "married"
newdata3[newdata3$marriage==2, 'marriage'] <- "single"

newdata3[newdata3$afftype==1, 'afftype'] <- "bipolar II"
newdata3[newdata3$afftype==2, 'afftype'] <- "depression"
newdata3[newdata3$afftype==3, 'afftype'] <- "bipolar I"

ggplot(newdata3, 
       aes(x = marriage, 
           fill = afftype)) + 
  geom_bar(position = position_dodge(preserve = "single"))
```
Work x Depression

```{r}

library(ggplot2)

newdata3[newdata3$work==1, 'work'] <- "working or studying"
newdata3[newdata3$work==2, 'work'] <- "unemployed/sick leave/pension"

newdata3[newdata3$afftype==1, 'afftype'] <- "bipolar II"
newdata3[newdata3$afftype==2, 'afftype'] <- "depression"
newdata3[newdata3$afftype==3, 'afftype'] <- "bipolar I"

ggplot(newdata3, 
       aes(x = work, 
           fill = afftype)) + 
  geom_bar(position = position_dodge(preserve = "single"))
```

## What do you not know how to do right now that you need to learn to answer your questions?

Machine learning.


## Do you plan on incorporating any machine learning techniques to answer your research questions? Explain.

I might. I will have to wait until next week to see if it's applicable. I am interested in learning and applying Nearest Neighbors Classification, K-Means Clustering


## Future Steps
I won't be adding new data, but I will be adding new research information to discuss about depression and gender. While women are more depressed than men, there are questions that came up, such as, how often are men questioned about depression symptoms, are the symptoms the same? Do men go to primary care or doctor visits in general as often as women? 




#References
